Activity Monitoring Analysis
========================================================

This R Markdown document is an analysis of the activity monitoring data.

Load the data
```{r echo=TRUE}
data <- read.csv('activity.csv')
```

## Mean total number of steps taken per day
### Histogram of the number of steps
```{r echo=TRUE}
data_clean <- na.omit(data)
summed <- aggregate(data_clean$steps, by = list(Date=data_clean$date), sum)
hist(summed$x, xlab = "Steps", main = "Histogram of total number of steps per day")
summary(summed)
```

### Mean & Median of the total number of steps per day
```{r echo=TRUE}
mean(summed$x)
median(summed$x)
```

## Average daily activity pattern
### Time series plot for average activity per interval
```{r echo=TRUE}
require(lattice)
avginterval <- aggregate(data_clean$steps, by = list(Interval=data_clean$interval), mean)
xyplot(x ~ Interval, data = avginterval, type = "l", main = "Average activity per interval", ylab = "Average Steps")
```
### Max 5-minute interval across all days
``` {r echo=TRUE}
avginterval[which.max(avginterval$x),1]
```

## Imputing Missing Data
### Total number of missing values
```{r echo=TRUE}
sum(!complete.cases(data))
```

### Filling missing data
Impute the missing steps by the mean of the steps for that interval across all days.
```{r echo=TRUE}
require(plyr)
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
data_filled <- ddply(data, ~interval, transform, steps = impute.mean(steps))
data_filled <- data_filled[order(data_filled$date), ]
```

### Histogram of total steps each day for imputed data
```{r echo=TRUE}
summed_filled <- aggregate(data_filled$steps, by = list(Date=data_filled$date), sum)
hist(summed_filled$x, xlab = "Steps", main = "Histogram of total number of steps per day")
summary(summed_filled)
```
### Mean & Median of the total number of steps per day for imputed data
```{r echo=TRUE}
mean(summed_filled$x)
median(summed_filled$x)
```

## Difference in activity patterns between weekdays and weekends
### Create weekday and weekend factors
``` {r echo=TRUE}
data_filled$weekdays <- weekdays(as.Date(data_filled$date))
data_filled$weekdays[data_filled$weekdays %in% c("Saturday", "Sunday")] <- "weekend"
data_filled$weekdays[!data_filled$weekdays %in% c("weekend")] <- "weekday"
data_filled$weekdays <- as.factor(data_filled$weekdays)
str(data_filled)
```

### Time series plot for average activity
``` {r echo=TRUE}
require(lattice)
steps_by_weekdays <- aggregate(data_filled$steps, by = list(weekdays=data_filled$weekdays, interval=data_filled$interval), mean)
xyplot(x ~ interval | weekdays, type = "l", data = steps_by_weekdays, xlab = "Interval", ylab = "Number of steps", layout = c(1, 2))
```